---
title: "R Notebook"
output: html_notebook
---

ctrl+alt+I to create chunk

ctrl+enter to run line

ctrl+shift+enter to run chunk

```{r}
x <- 'hello' #an expression, assignment
x      # auto print 
```

```{r}
print(x) #explicit print
```

```{r}
1:10 #sequence return numeric vector
```

```{r}
class(1:10)
```

```{r}
seq(from=1, to=20, by=2)
```

### Assignment:
 <- IS an assignment operator
 
 =  IS NOT (although it will work)
 
### Special numbers:

Inf

NaN (not a number)

NA  (not available)

 
```{r}
print(1/ 0)
print(1/Inf)
print(Inf*-Inf)
print(Inf/Inf)
print(0/0)
print(Inf*0)
```

NaN is numeric

NA values can have any class

NaN is also NA

NA is not NaN

```{r}
x <- c(NaN, NA)
print(is.nan(x))
print(is.na(x))
```


### Vectors

contain values of the same class


```{r}
y <- 25
class(y)
is.vector(y)
```

```{r}
y <- c(1:6, 5:1)
class(y)
```

```{r}
y <- c()
y
is.vector(y)
```

```{r}
y <- logical(length = 5)
y
class(y)
```


### Coercion 

automatic and explicit

```{r}
c(T, FALSE, 5)
c(TRUE, "FALSE", 5)
```

Mixture with character will be coerced to character

Mixture of numeric and logical will be coerced to numeric

```{r}
as.numeric(c('1', 2, 3))
as.logical(c(1, 0, 5, T))
as.character(c(T, F))
as.numeric(c('a', 'b'))
```

