---
title: "R Notation"
output: html_notebook
---

```{r}
deck <- read.csv('deck.csv')
deck
```

First, you’ll want to reshuffle the deck from time to time. And next, you’ll want to deal cards from the deck (one card at a time, whatever card is on top—we’re not cheaters).

To do these things, you’ll need to work with the individual values inside your data frame, a task essential to data science. For example, to deal a card from the top of your deck, you’ll need to write a function that selects the first row of values in your data frame, like this

Selecting values

```{r}
deck[ , ]
```

You have a choice when it comes to writing indexes. There are six different ways to write an index for R, and each does something slightly different. They are all very simple and quite handy, so let’s take a look at each of them. You can create indexes with:

* Positive integers
* Negative integers
* Zero
* Blank spaces
* Logical values
* Names

### Positive integers

R treats positive integers just like ij notation in linear algebra: deck[i,j] will return the value of deck that is in the ith row and the jth column.

```{r}
head(deck)
deck[1, 2]
class(deck[1, c(1, 2, 3)])
```

```{r}
new <- deck[1, c(1, 2, 3)]
deck
new
```

The same with vectors

```{r}
vec <- c(6, 1, 3, 6, 10, 5)
vec[1:3]
```

If you select two or more columns from a data frame, R will return a new data frame:

```{r}
deck[1:2, 1:2]
```

However, if you select a single column, R will return a vector:

```{r}
class(deck[1:2, 1])
```

If you would prefer a data frame instead, you can add the optional argument drop = FALSE between the brackets:

```{r}
deck[1:2, 1, drop = FALSE]
```

### Negative integer

Negative integers do the exact opposite of positive integers when indexing

```{r}
deck[-(3:52), 1:3]
```


```{r}
deck[-(2:52), 1:3]
```

```{r}
deck[c(-1, 1), 1]
```

### ZERO

R will return nothing from a dimension when you use zero as an index. This creates an empty object:

```{r}
class(deck[0, 0])
v <- c(1, 2, 3)
v[0]
```

### Blank spaces

You can use a blank space to tell R to extract every value in a dimension. 

```{r}
deck[ ,1, drop = FALSE]
```


### Logical values

If you supply a vector of TRUEs and FALSEs as your index, R will match each TRUE and FALSE to a row in your data frame (or a column depending on where you place the index). R will then return each row that corresponds to a TRUE

```{r}
deck[1, c(TRUE, TRUE, FALSE)]
```

```{r}
rows <- c(TRUE, F, F, F, F, F, F, F, F, F, F, F, F, F, F, F, 
  F, F, F, F, F, F, F, F, F, F, F, F, F, F, F, F, F, F, F, F, F, F, 
  F, F, F, F, F, F, F, F, F, F, F, F, F, T)
deck[rows, ]
```

### Names

Finally, you can ask for the elements you want by name—if your object has names (see Names). This is a common way to extract the columns of a data frame, since columns almost always have names

```{r}
deck[1, c("face", "suit", "value")]
```

```{r}
deck[ , "value"]
```

Deal and shuffle deck functions:

```{r}
shuffle <- function(cards) { 
  random <- sample(1:52, size = 52)
  cards[random, ]
}
```

```{r}
deal <- function(cards) {
  shuffle(cards)[1, ]
}
```

```{r}
deal(deck)
```


```{r}
shuffle(deck)
```

### Dollar sighns and double brackets

Two types of object in R obey an optional second system of notation. You can extract values from data frames and lists with the $ syntax. You will encounter the $ syntax again and again as an R programmer, so let’s examine how it works.

To select a column from a data frame, write the data frame’s name and the column name separated by a $. Notice that no quotes should go around the column name:

```{r}
deck$value
```

You can use the same $ notation with the elements of a list, if they have names. This notation has an advantage with lists, too. If you subset a list in the usual way, R will return a new list that has the elements you requested. This is true even if you only request a single element.

```{r}
lst <- list(numbers = c(1, 2), logical = TRUE, strings = c("a", "b", "c"))
lst
```

```{r}
lst[1]
sum(lst[1])
```

```{r}
lst$numbers
sum(lst$numbers)
```

If the elements in your list do not have names (or you do not wish to use the names), you can use two brackets, instead of one, to subset the list. This notation will do the same thing as the $ notation:

```{r}
lst[[3]]
```

